Networked Character Devices

### 2011-12-11 report/paper.tex
Minor formatting echnages, minor edits, and a few figures added

\usepackage[text={6in, 9in}, centering]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}

@@

\lstloadlanguages{C}
\lstset{
  language=C,
  basicstyle=\tiny,
  numbers=none,
  numberstyle=\footnotesize,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  frame=none,
  frameround=tttt,
  float
}

@@

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{arch-03-g.png}
  \end{center}
  \caption{The basic client-server model of NCD.}
  \label{fig:architecture}
\end{figure}

@@

Devices which have been imported on the client reside within the
devtmpfs filesystem. This is almost always mounted on \texttt{/dev}. By
default an imported device will take on the name
\texttt{/dev/netchar/importN}, where \texttt{N} is the next available
number. In the future (and without much effort), imported devices will
be able to take on their original name from the server. For example, an
input device can be located at \texttt{/dev/input/eventN}. Furthermore,
the user or administrator may choose any mapping (s)he see fits.

@@

The server is, as mentioned, a typical user-space daemon. It
currently only services a single file. Whether or not the server should
handle multiple exports is still up for evaluation.

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{proto-g.png}
  \end{center}
  \caption{NCD operates over a simple request reply protocol. Bulk data
  immediately follows formal messages, as required.}
  \label{fig:protocol}
\end{figure}

@@

Some of the basic features of the NCD kernel module are listed in
listing \ref{lst:client}. Firstly, NCD dynamically requests a major and
minor number from the running kernel \footnote{Statically assigned
major/minor numbers are no longer required and are discouraged for new
modules.}. These numbers are stored in a \texttt{struct dev\_t}. Once
these numbers are assigned, the kernel will pass all file operations on
character devices with the NCD major/minor number to the NCD module.

@@

Most importantly, \texttt{struct cdev} ties together the major/minor
number with the given set of file operations. With this it is possible
for a kernel module to register for many major/minor numbers (typically
a contiguous range), while having fine-grained control over which
subsets are handled by which distinct groups of file operations. This
will be important in the future as specific subsystems are handled
differently by NCD to increase functionality.

@@

A lot of magic is performed on character devices in userspace by
\texttt{udev}. Currently we are using \texttt{udev} to manage the
permissions of imported devices (listing \ref{lst:udevrule}).

@@

The Network Character Device system is still a work in
progress. We will need to Expand the system beyond it's current state
before the NCD concept is ready for production-level use.

Non the less, we have gained many insights into the advantages such a
system might provide. We have also encountered the difficulties
implementing such a system into an existing OS like Linux entails.

@@

As of this writing, we have completed a partial implementation of
the Network Character Device system. Current functionality includes:

\begin{packed_item}
\item Support for exporting a single character device from the server
\item Support for importing a single character device on the client
\item Client-side kernel-space Linux module implementation
\item Server-side user-space background process implementation
\item Support for open, close, read, and write calls
\item Basic Linux udev support for automatic node creation on client
\end{packed_item}

This set of core functionality provides a working proof-of-concept
Network Character Device system. This system fully supports the
export of a single character device from server to client over the
network and demonstrates the basic use and utility of such a system.

@@

\begin{packed_item}
\item Multi-device, multi-server client-side import support
\item Multi-device server-side export support
\item Support for exclusive use and protection of exported devices
\item Support for ioctl calls
\item Support for providing and obtaining metadata for exported
  devices
\item Support for ``advertisement'' of available exported devices
\item Support for integrating imported devices into other kernel
  subsystems (human interface subsystem, audio subsystem, etc)
\item Addition of ``NCD-admin'' utilities for managing and administering
  NCD system
\end{packed_item}

>>>>

\usepackage[text={6.5in, 9in}, centering]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}

@@

\lstloadlanguages{C}
\lstset{
  language=C,
  basicstyle=\tiny,
  numbers=none,
  numberstyle=\footnotesize,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  frame=none,
  frameround=tttt
}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{system-full-g.png}
  \caption{The NCD system. See Figure \ref{fig:system-working}
    for currently implemented system state.}
  \label{fig:system-full}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{arch-03-g.png}
  \caption{The basic client-server model of NCD.}
  \label{fig:architecture}
\end{figure}

@@

Devices which have been imported on the client reside within the
devtmpfs filesystem. This is almost always mounted on \texttt{/dev}. By
default an imported device will take on the name
\texttt{/dev/netchar/importN}, where \texttt{N} is the next available
number. In the future (and without much effort), imported devices will
be able to take on their original name from the server. For example, an
input device should be located at \texttt{/dev/input/eventN}. Furthermore,
the user or administrator may choose any mapping (s)he see fits.

@@

The server is a typical user-space daemon. It
currently only services a single file. Whether or not the server should
handle multiple exports is still up for evaluation.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{proto-g.png}
  \caption{NCD operates over a simple request reply protocol. Bulk data
    immediately follows formal messages, as required.}
  \label{fig:protocol}
\end{figure}

@@

Some of the basic features of the NCD kernel module are listed in
listing \ref{lst:client}. NCD dynamically requests a major and
minor number from the running kernel \footnote{Statically assigned
major/minor numbers are no longer required and are discouraged for new
modules.}. These numbers are stored in a \texttt{struct dev\_t}. Once
these numbers are assigned, the kernel will pass all file operations on
character devices with the NCD major/minor number to the NCD module.

@@

Most importantly, \texttt{struct cdev} ties together the major/minor
number with the given set of file operations. With this it is possible
for a kernel module to register many major/minor numbers (typically
a contiguous range), while having fine-grained control over which
subsets are handled by which distinct groups of file operations. This
will be important in the future as specific subsystems are handled
differently by NCD to increase functionality.

@@

A lot of magic is performed on character devices in userspace by
the Linux \texttt{udev} system.
Currently we are using \texttt{udev} to manage the
permissions of imported devices (listing \ref{lst:udevrule}).

@@

The Network Character Devices system is still a work in
progress. We will need to expand the system beyond it's current state
before the NCD concept is ready for production-level use.

None the less, we have gained many insights into the advantages such a
system might provide. We have also encountered the difficulties
implementing such a system into an existing OS like Linux entails.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{system-working-g.png}
  \caption{The NCD system with currently functioning portions
    highlighted.}
  \label{fig:system-working}
\end{figure}

As of this writing, we have completed a partial implementation of
the Network Character Devices system. Current functionality includes:

\begin{packed_item}
\item Support for exporting a single character device from the server
\item Support for importing a single character device on the client
\item Support for \texttt{open()}, \texttt{close()}, \texttt{read()},
  and \texttt{write()} calls
\item Basic Linux \texttt{udev} support for automatic node
  provisioning on the client
\end{packed_item}

This set of core functionality provides a working proof-of-concept
Network Character Devices system. This system fully supports the
export of a single character device from server to client over the
network and demonstrates the basic use and utility of such a system.

@@

\begin{packed_item}
\item Multi-device, multi-server client-side import support
\item Multi-device server-side export support
\item Support for exclusive use and protection of exported devices
\item Support for ioctl calls
\item Support for providing and obtaining metadata for exported
  devices
\item Support for ``advertisement'' of available exported devices
\item Support for integrating imported devices into other kernel
  subsystems (human interface subsystem, audio subsystem, etc)
\item Addition of UDP and TCP support with user selection
\item Addition of graceful connect and disconnect handling in real-time
\end{packed_item}
